---
title: "Visualizing Palmer Penguins (Exploratory Data Analysis)"
output: github_document
---

```{r load-packages, include=FALSE}
library(tidyverse)
library(palmerpenguins)
library(ggplot2)
```

# Palmer Penguins

### Scatterplot

``` {r scatterplot}
penguin_plot = ggplot(data = penguins, # gives a blank canvas
       mapping = aes(x = bill_depth_mm, # gives x variables (can be histogram)
                     y = bill_length_mm, # makes y relationship
                     colour = species)) + # specifies a color difference
  geom_point() + # plots points
  labs(title = "Bill depth and length", # adds a title
       subtitle = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
       x = "Bill depth (mm)", y = "Bill length (mm)", # make column names logical
       colour = "Species", # specify what to call the color legend
       caption = "Source: Palmer Station LTER / palmerpenguins package") + # give some caption to make it professional
  scale_colour_viridis_d()

penguin_plot
```

- aes() - every visualization needs aesthetics (or else you can't see it)
  - colour, shape, size, alpha (transparency)
  - bar plot --> fill is different from color (color is outside, fill is inside)
  - scatterplot --> color is intuitive
  
### Adding Shapes to the scatterplot
``` {r shape-scatterplot}
# use previous data
shape_plot = 
ggplot(penguins,
       aes(x = bill_depth_mm, 
           y = bill_length_mm,
           colour = species,
           shape = island)) +
  geom_point() +
  scale_colour_viridis_d()

shape_plot
```

### Faceting
```{r faceting}
facet_plot = ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + 
  geom_point() +
  facet_grid(species ~ island)

facet_plot
```

### Facet and Color
```{r facet-color}
ggplot(
  penguins, 
  aes(x = bill_depth_mm, 
      y = bill_length_mm, 
      color = island)) +
  geom_point() +
  facet_grid(species ~ sex) +
  scale_color_viridis_d()
```


### Facet Wrap
```{r facet-wrap}
ggplot(penguins, aes(x = bill_depth_mm, y = bill_length_mm)) + 
  geom_point() +
  facet_wrap(~ species, ncol = 2)
```
- good for comparing trends in different variables
- not overlaid anymore




