---
title: "Working with Numerical Data"
output: github_document
---

```{r load-packages, include=FALSE}
library(tidyverse)
library(openintro)
```

### Check out the data!
```{r data-variables}
library(openintro)
glimpse(loans_full_schema)
```

### Make a loans object
```{r object}
loans <- loans_full_schema %>%
  select(loan_amount, interest_rate, term, grade, 
         state, annual_income, homeownership, debt_to_income)
glimpse(loans)
```

### Visualizing Distributions - Histogram
```{r histogram}
basic_histogram = ggplot(loans, aes(x = loan_amount)) + # add the data as an aesthetic
  geom_histogram(binwidth = 5000)  # does a ton under the hood --> does frequency count

basic_histogram
```
- can leave the histogram parameter blank for it to autochoose

### Customizing Histograms
```{r customized-histogram}
basic_histogram + labs(
    x = "Loan amount ($)",
    y = "Frequency",
    title = "Amounts of Lending Club loans"
  )
```

## Make it look pretty!
```{r pretty-histogram}
pretty_histogram = ggplot(loans, aes(x = loan_amount, 
                  fill = homeownership)) +
  geom_histogram(binwidth = 5000,
                 alpha = 0.5) +
  labs(
    x = "Loan amount ($)",
    y = "Frequency",
    title = "Amounts of Lending Club loans"
  )

pretty_histogram
```

### Facets?
```{r facet-histogram}
pretty_histogram +
  facet_wrap(~ homeownership, nrow = 3)
```

- this is usually a bit better for histograms

### Density
```{r density}
density_plot = ggplot(loans, aes(x = loan_amount)) +
  geom_density(adjust = 2)

density_plot
```

### Add Some Labels
```{r labeled-density}
labeled_density_plot = density_plot + labs(
    x = "Loan amount ($)",
    y = "Density",
    title = "Amounts of Lending Club loans"
  )

labeled_density_plot
```

### Add a categorical variable
```{r categorical-density}
ggplot(loans, aes(x = loan_amount, 
                  fill = homeownership)) +
  geom_density(adjust = 2, 
               alpha = 0.5) +
  labs(
    x = "Loan amount ($)",
    y = "Density",
    title = "Amounts of Lending Club loans", 
    fill = "Homeownership"
  )
```

### Box Plots
```{r boxplot}
ggplot(loans, aes(x = interest_rate)) +
  geom_boxplot()
```

### Custom Box Plots
```{r custom-box-plots}
ggplot(loans, aes(x = interest_rate,
                  y = grade)) +
  geom_boxplot() +
  labs(
    x = "Interest rate (%)",
    y = "Grade",
    title = "Interest rates of Lending Club loans",
    subtitle = "by grade of loan"
  )
```


### Bar Plots
```{r bar-plot}
basic_bar = ggplot(loans, aes(x = homeownership)) +
  geom_bar() # sorting categorical data

basic_bar
```

### Segmented Bar Plots
```{r segmented-bar-plot}
ggplot(loans, aes(x = homeownership, # segments on x-axis
                  fill = grade)) + # changes fill based on categorical variable grade
  geom_bar()
```
- shows how many of each